Enhancing Forecast Reconciliation: A Study of Alternative Covariance Estimators

### 2025-10-21 thesis/paper_draft.qmd
fixing the font size and spacing, minor fix to scale_fill_manual for bar plots

---
title: "Enhancing Forecast Reconciliation: A Study of Alternative Covariance Estimators"
author: 
  - Vincent Su
  - Shanika Wickramasuriya (supv.)
  - George Athanasopoulos (supv.)
format: 
  pdf:
    include-in-header:
      text: |
        <!-- \usepackage{amsmath} -->
        <!-- \usepackage{mathspec} -->
        \usepackage{algorithm}
        \usepackage{algpseudocode}
        \usepackage[colorlinks=true, linkcolor=blue, citecolor=blue, urlcolor=blue]{hyperref}
        \usepackage{array}
        \usepackage{float}
        \usepackage{xcolor}
    mathspec: true
    number-sections: true
# crossref:
#   custom:
#     - kind: float
#       key: alg
#       reference-prefix: Algorithm
#       caption-prefix: Algorithm
#       # for PDF output only, tell LaTeX to use the `algorithm` env:
#       latex-env: algorithm
link-citations: true
execute:
  cache: true
  echo: false
  warning: false
bibliography: D:/Github/Recon_Honours_Thesis/references.bib
# bibliographystyle: apa
csl: D:/Github/Recon_Honours_Thesis/apa.csl
editor_options: 
  chunk_output_type: console
---

@@

@@

p_c95_lvl <- tourism_coverage95 |> 
  group_by(level, .model) |> 
  summarise(mean_coverage = mean(coverage)) |> 
  encode_recon_models() |> 
  ggplot(aes(x = model_label, y = mean_coverage, fill = family, alpha = pc_k)) +
    geom_col(color = "black") +
    scale_alpha_manual(values = c("0" = 1, "1" = 0.7, "2" = 0.4), guide = "none") +
    labs(x = "", y = "Empirical Coverage",
         title = "Empirical coverage of 95% prediction intervals by level") +
    geom_hline(yintercept = 0.95, linetype = "dashed", color = "black") + # draw line for coverage
    facet_wrap(~ level, ncol=2, scales = "free_y") +
    recon_fill_scale +
    theme_minimal() +
    # rotate the x axis labels
    theme(axis.text.x = element_text(angle = 20, hjust = 0.5))

@@

This paper aims to address the limitations of the current Minimum Trace reconciliation method, a method that has garnered significant attention in both academic and practical forecasting contexts. We propose the NOVELIST estimator to address the lack of flexibility in the uniform shrinkage target of the traditional shrinkage estimator. Additionally, we introduce PC-adjusted variants of shrinkage and NOVELIST to take into account dominant latent factors, as well as multi-step covariance variants that relax the proportionality assumption.

Empirical results on Australian domestic tourism data reveal several insights. For point forecasts, shrinkage (and its PC-adjusted version) remains a robust default choice, while NOVELIST demonstrates improved performance for probabilistic forecasts. Adjusting for a single dominant principal component consistently enhances performance across both point and probabilistic forecasts, suggesting its utility when such latent structures are present. More complex adjustments, such as incorporating multiple principal components or horizon-specific covariances, tend to introduce additional estimation noise without notable benefits.

However, several limitations and avenues for future research remain. More intricate simulation studies are necessary to further highlight the differences between shrinkage and NOVELIST, particularly in scenarios involving factor models. The current implementation does not standardise data for PC-variant methods, which may lead to higher-level series largely influencing the principal components. Future work could also explore alternative methods that leverage cross-series information or eigenvalue/eigenvector structures. Additionally, understanding why MinT with shrinkage underperforms in the Winkler 95% score, especially at higher aggregation levels, needs further investigation. Finally, evaluating h-step-ahead probabilistic forecasts remains an open area for further research.

>>>>

---
title: "Enhancing Forecast Reconciliation: A Study of Alternative Covariance Estimators"
author: 
  - Vincent Su
  - Shanika Wickramasuriya (supv.)
  - George Athanasopoulos (supv.)
format: 
  pdf:
    fontsize: 11pt
    linestretch: 1.15
    include-in-header:
      text: |
        <!-- \usepackage{amsmath} -->
        <!-- \usepackage{mathspec} -->
        \usepackage{algorithm}
        \usepackage{algpseudocode}
        \usepackage[colorlinks=true, linkcolor=blue, citecolor=blue, urlcolor=blue]{hyperref}
        \usepackage{array}
        \usepackage{float}
        \usepackage{xcolor}
    mathspec: true
    number-sections: true
# crossref:
#   custom:
#     - kind: float
#       key: alg
#       reference-prefix: Algorithm
#       caption-prefix: Algorithm
#       # for PDF output only, tell LaTeX to use the `algorithm` env:
#       latex-env: algorithm
link-citations: true
execute:
  cache: true
  echo: false
  warning: false
bibliography: D:/Github/Recon_Honours_Thesis/references.bib
# bibliographystyle: apa
csl: D:/Github/Recon_Honours_Thesis/apa.csl
editor_options: 
  chunk_output_type: console
---

@@

\vspace{2cm}

@@

p_c95_lvl <- tourism_coverage95 |> 
  group_by(level, .model) |> 
  summarise(mean_coverage = mean(coverage)) |> 
  encode_recon_models() |> 
  ggplot(aes(x = model_label, y = mean_coverage, fill = family, alpha = pc_k)) +
    geom_col(color = "black") +
    scale_alpha_manual(values = c("0" = 1, "1" = 0.7, "2" = 0.4), guide = "none") +
    labs(x = "", y = "Empirical Coverage",
         title = "Empirical coverage of 95% prediction intervals by level") +
    geom_hline(yintercept = 0.95, linetype = "dashed", color = "black") + # draw line for coverage
    facet_wrap(~ level, ncol=2, scales = "free_y") +
    recon_fill_scale_w_base +
    theme_minimal() +
    # rotate the x axis labels
    theme(axis.text.x = element_text(angle = 20, hjust = 0.5))

@@

This paper tackles a central bottleneck in Minimum Trace (MinT) reconciliation: obtaining reliable, positive-definite estimates of the base-forecast error covariance. The original practice, following @Wickramasuriya2019-xq, estimates the one-step covariance via diagonal-target shrinkage and obtains multi-step horizons by proportional scaling. We identified three limitations of this approach--uniform shrinkage, lack of factor awareness, and horizon-invariant dependence--and proposed alternatives that address them individually and in combination. Specifically, we adopt NOVELIST to introduce a more flexible shrinkage target; develop PC-adjusted variants of shrinkage and NOVELIST to preserve dominant latent components in the data; and consider multi-step covariance constructions that relax proportional scaling assumption.

Empirical evidence from Australian domestic tourism yields several insights. For point reconciliation, shrinkage (and its PC-adjusted version) remains a robust default choice, while NOVELIST demonstrates improved performance for probabilistic forecasts. Across both settings, adjusting for a single dominant principal component consistently enhances performance, indicating its utility when such latent structures are present. More complex adjustments, such as incorporating multiple principal components or horizon-specific covariances, tend to introduce additional idiosyncratic noise without notable benefits.

Several limitations and directions for future work remain. First, more intricate simulation designs are needed to further highlight the differences between shrinkage and NOVELIST, particularly under explicit factor models and varying sparsity/noise levels. Second, our current PC-adjusted implementations do not standardise series prior to factor extraction, which may lead to higher-level series largely influencing the principal components. Third, we can explore alternative methods that leverage cross-series information or eigenvalue/eigenvector structures. Fourth, the observed inferior performance of MinT with shrinkage under the Winkler 95% score at higher aggregation levels needs detailed investigation. Finally, evaluating genuinely $h$-step probabilistic reconciliation remains an open avenue, including how to best regularise multi-step covariances given limited effective samples.
